\pagestyle{plain}
\chapter{Acknoledgements}
\setcounter{page}{1}
I first would like to apologies to all the people I will possibly forget in this section. 

To my parents, who always pushed me and forced me go beyond the limits of what I thought was possible.

To my marvelous girlfriend, Nerea Castedo, for her love, for her support, for being always here when I needed her and for making me dream of a bright and common future.
To her parents and her family, who accepted me as one of their own, making me discover a new world and culture unbeknownst to me before meeting them.

To my friends Shehab Zaineldine, Jorge Rodríguez, Álvaro López, Ionut Sorin, Alba Ochoa, Javier López and François Delattre who were always here to cheer me up even when the time were rough with a special mention to Manuel Rodriguez Gonzalo that was without a doubt the most reliable and trustful workmate and friend I ever had.

To the teachers of the University Carlos III de Madrid without whom I wouldn't even hope to realize this project. Special mention to the ARCOS department who gave me the taste of low level programming.

To Francisco Javier Blas, who helped me in many different ways throughout this whole degree.

To Luis Cantarero that was able to help me throughout the numerous administrative problems that I encountered in my student life in the UC3M.

To my tutor, Javier Fernández Muñoz, who trusted me with this project and helped me throughout its realization.

José Manuel Peso, Mánel, Roberto León and Javier Espinosa for the wisdom and knowledge that you shared with me during all these months.

To all the people that are not named in this section but who are in my heart.

Thanks to all of you, for your trust, support, help and time. I hope to deserve everything you offered me and not to disappoint you.
